---
title: "RVE Manuscript WT v ztl Half-Life for RVE8"
author: "Cassandra"
date: "2022-11-18"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Background

RVE8 protein degradation was assessed at ZT7 and ZT19 in monochromatic blue, monochromatic red, or constant dark light conditions. This was done both in a WT background and a *ztl* mutant background.       

Stacey suggested comparing the half-life of RVE8 between the WT and *ztl* backgrounds in constant blue light at a single time point (ZT7 or ZT19).    

# Summary

At ZT7 in constant blue light, there is a significant difference between RVE8 deg rate (but not half-life) in WT background and *ztl* background.    
At ZT19 in constant blue light, there is no significant difference between RVE8 deg rate (or half-life) in WT background and *ztl* background.     

```{r setup}
# for graphing
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())

# for Bayesian analysis 
library(posterior)
library(brms)
```

```{r}
#setwd("C:/Users/contr/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")

#setwd("F:/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")
```

# RVE8 WT vs ztl  

## load and check data

```{r}
# load data
wt <- read_csv("CHX_RVE8_WT.csv")
ztl <- read_csv("CHX_RVE8_ztl.csv")

# join data frames together
rve8 <- full_join(wt, ztl)

# rename columns
colnames(rve8) <- c("exp", "blot", "label", "time", "zt", "light", "gt", "rep", "target", "abundance", "scaled", "zt_scaled")

# set factors, remove unnecessary columns
# also add exp_rep column to separate bio reps 
rve8 <- rve8 %>% 
  select(-c(blot, label)) %>% 
  mutate(exp = as_factor(exp),
         zt = as_factor(zt),
         light = as_factor(light),
         gt = as_factor(gt),
         target = as_factor(target),
         exp_rep = as_factor(paste0(exp, "_", rep)))

summary(rve8)
```

```{r}
rve8 %>% 
  ggplot(aes(time, zt_scaled, color = gt)) +
  geom_point() + 
  geom_line(aes(linetype = exp_rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)

rve8 %>% 
  ggplot(aes(time, scaled, color = gt)) +
  geom_point() + 
  geom_line(aes(linetype = exp_rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)
```

```{r}
# set up data subsets for each color
rve8b <- rve8 %>% filter(light == "60B")

# red subsets
rve8r <- rve8 %>% filter(light == "60R")

# dark subsets
rve8d <- rve8 %>% filter(light == "DD")
```

```{r}
rve8b %>%
  ggplot(aes(x=time, y=zt_scaled, color=gt)) +
  geom_smooth() +
  scale_y_log10() +
  facet_wrap(~zt) +
  ggtitle("blue")
```

```{r}
rve8r %>%
  ggplot(aes(x=time, y=zt_scaled, color=gt)) +
  geom_smooth() +
  scale_y_log10() +
  facet_wrap(~zt) +
  ggtitle("red")
```

```{r}
rve8d %>%
  ggplot(aes(x=time, y=zt_scaled, color=gt)) +
  geom_smooth() +
  scale_y_log10() +
  facet_wrap(~zt) +
  ggtitle("dark")
```
## Blue


#### fit and compare models

```{r}
# first allow k and intercept to vary by gt and zt with random effect per rep 
m1 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k + N0 ~ gt * zt +  (1|exp_rep),
           nl = TRUE)


# assign priors
priors1 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "N0"),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit1 <- brm(formula = m1,
               prior = priors1,
               data = rve8b,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.95))
```

```{r}
# check model 
summary(fit1)

plot(fit1, ask = FALSE)
```

Rhat = 1, so parameter space was sampled well. Diagnostic plots look okay.   

```{r}
# no random effect for N0
m2 <- bf(zt_scaled ~ N0 * exp(-k*time),
         k ~ gt * zt +  (1|exp_rep),
         N0 ~ gt * zt,
         nl = TRUE)


# assign priors
priors2 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit2 <- brm(formula = m2,
               prior = priors2,
               data = rve8b,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
# check model 
summary(fit2)

plot(fit2, ask = FALSE)
```

Rhat = 1, plots look okay.    

```{r}
# no random effect for N0 or k
m3 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k + N0 ~ gt * zt,
           nl = TRUE)


# assign priors
priors3 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma))

# fit model 
fit3 <- brm(formula = m3,
               prior = priors3,
               data = rve8b,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.95))
```

```{r}
# check model 
summary(fit3)

plot(fit3, ask = FALSE)
```

Rhat = 1, plots look okay.   

```{r}
# no zt effects on N0
m4 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ gt * zt +  (1|exp_rep),
           N0 ~ gt + (1|exp_rep), 
           nl = TRUE)


# assign priors
priors4 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "N0"),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit4 <- brm(formula = m4,
               prior = priors4,
               data = rve8b,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
summary(fit4)
plot(fit4, ask=FALSE)
```
```{r}
# no zt effects and no random effects on N0
m5 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ gt * zt +  (1|exp_rep),
           N0 ~ gt, 
           nl = TRUE)


# assign priors
priors5 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit5 <- brm(formula = m5,
               prior = priors5,
               data = rve8b,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
summary(fit5)
plot(fit5, ask = FALSE)
```

```{r}
# no zt or gt effects on N0
m6 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ gt * zt +  (1|exp_rep),
           N0 ~ 1 + (1|exp_rep), 
           nl = TRUE)


# assign priors
priors6 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "N0"),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit6 <- brm(formula = m6,
               prior = priors6,
               data = rve8b,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```
```{r}
summary(fit6)
plot(fit6, ask=FALSE)
```

```{r}
# no zt or gt effects on N0; no random effects
m7 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ gt * zt,
           N0 ~ 1, 
           nl = TRUE)


# assign priors
priors7 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma))

# fit model 
fit7 <- brm(formula = m7,
               prior = priors7,
               data = rve8b,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
summary(fit7)
plot(fit7, ask=FALSE)
```

```{r}
# compare the  models
LOO(fit1, fit2, fit3, fit4, fit5, fit6, fit7)
```

Many are roughly equivalent.  Let's go with the relatively simple model 6

Resample to sample prior, needed for hypothesis testing, below
```{r}
fit6 <- update(fit6,    
               sample_prior = "yes",
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99) )
```

```{r}
summary(fit6)
plot(fit6, ask=FALSE)
```


### compare deg rate between ztl and WT

```{r}
# is ztl deg rate significantly different from WT deg rate at zt19?
hyp1 <- hypothesis(fit6, hypothesis = "k_gtztl + k_gtztl:zt19 = 0")

hyp1

#plot(hyp1)
```
No difference

Rate at ZT 7
```{r}
# is ztl deg rate significantly different from WT deg rate at zt7?
hyp2 <- hypothesis(fit6, hypothesis = "k_gtztl = 0")

hyp2

#plot(hyp2)
```
