---
title: "RVE Manuscript RvB Half-Life Interaction"
author: "Julin and Cassandra"
date: "2023-07-23"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Background

RVE8 protein degradation was assessed at ZT7 and ZT19 in monochromatic blue, monochromatic red, or constant dark light conditions.     

Stacey suggested comparing the half-life of RVE8 in constant blue vs constant red light at a single time point (ZT7 or ZT19).   
Probably want to do this in both WT and ztl backgrounds. Should this also be done for RVE4 at ZT5?

Julin thinks this may work better as an interaction model (ZT * light).  That is what is being tried here.

# Summary

  

```{r setup}
# for graphing
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())

# for Bayesian analysis 
library(posterior)
library(brms)
```

```{r}
#setwd("C:/Users/contr/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")

#setwd("F:/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")
```

# RVE8 WT background 

## load and check data

```{r}
# load data
rve8 <- read_csv("CHX_RVE8_WT.csv")

# rename columns
colnames(rve8) <- c("exp", "blot", "label", "time", "zt", "light", "gt", "rep", "target", "abundance", "scaled", "zt_scaled")

# set factors, remove unnecessary columns
# also add exp_rep column to separate bio reps 
rve8 <- rve8 %>% 
  select(-c(blot, label, gt)) %>% 
  mutate(exp = as_factor(exp),
         zt = as_factor(zt),
         light = as_factor(light),
         target = as_factor(target),
         exp_rep = as_factor(paste0(exp, "_", rep)))

summary(rve8)
```

```{r}
rve8 %>% 
  ggplot(aes(time, zt_scaled, color = exp_rep)) +
  geom_point() + 
  geom_line(aes(linetype = exp_rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)

rve8 %>% 
  ggplot(aes(time, scaled, color = exp_rep)) +
  geom_point() + 
  geom_line(aes(linetype = exp_rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)
```
make a plot with just red vs blue:

```{r}
rve8 %>% 
  filter(light != "DD", zt=="19") %>%
  ggplot(aes(time, zt_scaled, color = light, group=str_c(light,exp_rep))) +
  scale_color_manual(values=c("60B"="blue", "60R"="red")) +
  geom_point() + 
  geom_line() +
  scale_y_log10()
```

```{r}
rve8 %>% 
  filter(light != "DD", zt=="19") %>%
  ggplot(aes(time, zt_scaled, color = light)) +
  scale_color_manual(values=c("60B"="blue", "60R"="red")) +
  geom_smooth() +
  scale_y_log10()
```
```{r}
rve8 %>% 
  filter(light != "DD", zt=="7") %>%
  ggplot(aes(time, zt_scaled, color = light)) +
  scale_color_manual(values=c("60B"="blue", "60R"="red")) +
  geom_smooth() +
  scale_y_log10()
```

```{r}
# set up data subsets for no dark

rve8rb <- rve8 %>% filter(light != "DD") %>%
  mutate(light=relevel(light, ref="60R"))
```

## ZT7

### fit and compare models

Use non-linear formula and set priors similar to initial RVE8 analysis.   

```{r}
# first allow k and intercept to vary by light and zt with random effect per rep 
m1 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k + N0 ~ light * zt +  (1|exp_rep),
           nl = TRUE)


# assign priors
priors1 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "N0"),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit1 <- brm(formula = m1,
               prior = priors1,
               data = rve8rb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.95))
```

```{r}
# check model 
summary(fit1)

plot(fit1, ask = FALSE)
```

Rhat = 1, so parameter space was sampled well. Diagnostic plots look okay.   

```{r}
# no random effect for N0
m2 <- bf(zt_scaled ~ N0 * exp(-k*time),
         k ~ light * zt +  (1|exp_rep),
         N0 ~ light * zt,
         nl = TRUE)


# assign priors
priors2 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit2 <- brm(formula = m2,
               prior = priors2,
               data = rve8rb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
# check model 
summary(fit2)

plot(fit2, ask = FALSE)
```

Rhat = 1, plots look okay.    

```{r}
# no random effect for N0 or k
m3 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k + N0 ~ light * zt,
           nl = TRUE)


# assign priors
priors3 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma))

# fit model 
fit3 <- brm(formula = m3,
               prior = priors3,
               data = rve8rb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.95))
```

```{r}
# check model 
summary(fit3)

plot(fit3, ask = FALSE)
```

Rhat = 1, plots look okay.   

```{r}
# no zt effects on N0
m4 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ light * zt +  (1|exp_rep),
           N0 ~ light + (1|exp_rep), 
           nl = TRUE)


# assign priors
priors4 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "N0"),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit4 <- brm(formula = m4,
               prior = priors4,
               data = rve8rb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
summary(fit4)
plot(fit4, ask=FALSE)
```
```{r}
# no zt effects and no random effects on N0
m5 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ light * zt +  (1|exp_rep),
           N0 ~ light, 
           nl = TRUE)


# assign priors
priors5 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit5 <- brm(formula = m5,
               prior = priors5,
               data = rve8rb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
summary(fit5)
plot(fit5, ask = FALSE)
```

```{r}
# no zt or light effects on N0
m6 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ light * zt +  (1|exp_rep),
           N0 ~ 1 + (1|exp_rep), 
           nl = TRUE)


# assign priors
priors6 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "N0"),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit6 <- brm(formula = m6,
               prior = priors6,
               data = rve8rb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```
```{r}
summary(fit6)
plot(fit6, ask=FALSE)
```

```{r}
# no zt or light effects on N0; no random effects
m7 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ light * zt,
           N0 ~ 1, 
           nl = TRUE)


# assign priors
priors7 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma))

# fit model 
fit7 <- brm(formula = m7,
               prior = priors7,
               data = rve8rb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
summary(fit7)
plot(fit7, ask=FALSE)
```

```{r}
# compare the  models
LOO(fit1, fit2, fit3, fit4, fit5, fit6, fit7)
```

Many are roughly equivalent.  Let's go with model 6

Resample to sample prior, needed for hypothesis testing, below
```{r}
fit6 <- update(fit6,    
               sample_prior = "yes",
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99) )
```

```{r}
summary(fit6)
plot(fit6, ask=FALSE)
```


### compare deg rate between blue and red light

```{r}
# is red deg rate significantly different from blue deg rate at zt19?
hyp1 <- hypothesis(fit6, hypothesis = "k_light60B + k_light60B:zt19 > 0")

hyp1

#plot(hyp1)
```
blue is greater than red at ZT19

Rate at ZT 7
```{r}
# is red deg rate significantly different from blue deg rate at zt7?
hyp2 <- hypothesis(fit6, hypothesis = "k_light60B = 0")

hyp2

plot(hyp2)
```

### calculate half-lives and compare between blue and red at ZT19

```{r}
# first get coefficients
coefs <- fixef(fit6)[, "Estimate"]

#for blue
log(2) / sum(coefs[c("k_Intercept", "k_light60B", "k_zt19", "k_light60B:zt19")])

#for red
log(2) / sum(coefs[c("k_Intercept", "k_zt19")])
```

Half-lives are pretty consistent with previous calculations, nice.   

```{r}
# is red half-life significantly different from blue half-life?
hyp3 <- hypothesis(fit6, hypothesis = "(log(2) / (k_Intercept + k_zt19)) > (log(2) / (k_Intercept + k_light60B + k_zt19 + k_light60B:zt19))")
hyp3

plot(hyp3)
```

Red half-life is significantly lower at ZT19

### calculate half-lives and compare between blue and red at ZT7

```{r}
# first get coefficients
coefs <- fixef(fit6)[, "Estimate"]

#for blue
log(2) / sum(coefs[c("k_Intercept", "k_light60B")])

#for red
log(2) / sum(coefs[c("k_Intercept")])
```

```{r}
# is red half-life significantly different from blue half-life?
hyp4 <- hypothesis(fit6, hypothesis = "(log(2) / (k_Intercept)) = (log(2) / (k_Intercept + k_light60B))")
hyp4

plot(hyp4)
```

Red half-life is not different from blue at ZT7


# RVE8 ztl background 

## load and check data

```{r}
# load data
ztl <- read_csv("CHX_RVE8_ztl.csv")

# rename columns
colnames(ztl) <- c("exp", "blot", "label", "time", "zt", "light", "gt", "rep", "target", "abundance", "scaled", "zt_scaled")

# set factors, remove unnecessary columns
# also add exp_rep column to separate bio reps 
ztl <- ztl %>% 
  select(-c(blot, label)) %>% 
  mutate(exp = as_factor(exp),
         zt = as_factor(zt),
         light = as_factor(light),
         gt = as_factor(gt),
         target = as_factor(target),
         exp_rep = as_factor(paste0(exp, "_", rep)))

summary(ztl)
```

make a plot with just red vs blue:

```{r}
ztl %>% 
  filter(light != "DD", zt=="19") %>%
  ggplot(aes(time, zt_scaled, color = light, group=str_c(light,exp_rep))) +
  scale_color_manual(values=c("60B"="blue", "60R"="red")) +
  geom_point() + 
  geom_line() +
  scale_y_log10()
```

```{r}
ztl %>% 
  filter(light != "DD", zt=="19") %>%
  ggplot(aes(time, zt_scaled, color = light)) +
  scale_color_manual(values=c("60B"="blue", "60R"="red")) +
  geom_smooth() +
  scale_y_log10()
```

```{r}
ztl %>% 
  filter(light != "DD", zt=="7") %>%
  ggplot(aes(time, zt_scaled, color = light, group=str_c(light,exp_rep))) +
  scale_color_manual(values=c("60B"="blue", "60R"="red")) +
  geom_point() + 
  geom_line() +
  scale_y_log10()
```

```{r}
ztl %>% 
  filter(light != "DD", zt=="7") %>%
  ggplot(aes(time, zt_scaled, color = light)) +
  scale_color_manual(values=c("60B"="blue", "60R"="red")) +
  geom_smooth() +
  scale_y_log10()
```

```{r}
# set up data subsets for no dark

ztlrb <- ztl %>% filter(light != "DD") %>%
  mutate(light=relevel(light, ref="60R"))
```


### fit and compare models

Use non-linear formula and set priors similar to initial RVE8 analysis.  

Intercept looks to be different in R vs B for ZT7, so need to account for that in the model.


```{r}
# first allow k and intercept to vary by light and zt with random effect per rep 
m1 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k + N0 ~ light * zt +  (1|exp_rep),
           nl = TRUE)


# assign priors
priors1 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "N0"),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit1ztl <- brm(formula = m1,
               prior = priors1,
               data = ztlrb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
# check model 
summary(fit1ztl)

plot(fit1ztl, ask = FALSE)
```

Rhat = 1, so parameter space was sampled well. Diagnostic plots look okay.  1 Divergent transition. 

```{r}
# no random effect for N0
m2 <- bf(zt_scaled ~ N0 * exp(-k*time),
         k ~ light * zt +  (1|exp_rep),
         N0 ~ light * zt,
         nl = TRUE)


# assign priors
priors2 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit2ztl <- brm(formula = m2,
               prior = priors2,
               data = ztlrb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
# check model 
summary(fit2ztl)

plot(fit2ztl, ask = FALSE)
```

Rhat = 1, plots look okay.    

```{r}
# no random effect for N0 or k
m3 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k + N0 ~ light * zt,
           nl = TRUE)


# assign priors
priors3 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma))

# fit model 
fit3ztl <- brm(formula = m3,
               prior = priors3,
               data = ztlrb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.95))
```

```{r}
# check model 
summary(fit3ztl)

plot(fit3ztl, ask = FALSE)
```

Rhat = 1, plots look okay.   

```{r}
# # no zt effects on N0
# # does not sample well
# m4 <- bf(zt_scaled ~ N0 * exp(-k*time),
#            k ~ light * zt +  (1|exp_rep),
#            N0 ~ light + (1|exp_rep), 
#            nl = TRUE)
# 
# 
# # assign priors
# priors4 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
#              prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
#              prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
#              prior(cauchy(0, 1), class = sigma),
#              prior(cauchy(0, 1), class = sd, nlpar = "N0"),
#              prior(cauchy(0, 1), class = sd, nlpar = "k"))
# 
# # fit model 
# fit4ztl <- brm(formula = m4,
#                prior = priors4,
#                data = ztlrb,
#                cores = 4,
#                iter = 8000,
#                control = list(adapt_delta = 0.99))
```

```{r}
# summary(fit4ztl)
# plot(fit4ztl, ask=FALSE)
```

```{r}
# # no zt effects and no random effects on N0
## Does not sample well
# m5 <- bf(zt_scaled ~ N0 * exp(-k*time),
#            k ~ light * zt +  (1|exp_rep),
#            N0 ~ light, 
#            nl = TRUE)
# 
# 
# # assign priors
# priors5 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
#              prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
#              prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
#              prior(cauchy(0, 1), class = sigma),
#              prior(cauchy(0, 1), class = sd, nlpar = "k"))
# 
# # fit model 
# fit5ztl <- brm(formula = m5,
#                prior = priors5,
#                data = ztlrb,
#                cores = 4,
#                iter = 8000,
#                control = list(adapt_delta = 0.99))
```

```{r}
# summary(fit5ztl)
# plot(fit5ztl, ask = FALSE)
```

```{r}
# no zt or light effects on N0
m6 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ light * zt +  (1|exp_rep),
           N0 ~ 1 + (1|exp_rep), 
           nl = TRUE)


# assign priors
priors6 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "N0"),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
fit6ztl <- brm(formula = m6,
               prior = priors6,
               data = ztlrb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```
```{r}
summary(fit6ztl)
plot(fit6ztl, ask=FALSE)
```

```{r}
# no zt or light effects on N0; no random effects
m7 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k ~ light * zt,
           N0 ~ 1, 
           nl = TRUE)


# assign priors
priors7 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma))

# fit model 
fit7ztl <- brm(formula = m7,
               prior = priors7,
               data = ztlrb,
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99))
```

```{r}
summary(fit7ztl)
plot(fit7ztl, ask=FALSE)
```

#4 and 5 do not sample well, so not including them.
```{r}
# compare the  models
LOO(fit1ztl, fit2ztl, fit3ztl, fit6ztl, fit7ztl)
```

I want to use a model with an interaction for N0, so let's go with 3

Resample to sample prior, needed for hypothesis testing, below
```{r}
fit3ztl <- update(fit3ztl,    
               sample_prior = "yes",
               cores = 4,
               iter = 8000,
               control = list(adapt_delta = 0.99) )
```

```{r}
summary(fit3ztl)
plot(fit3ztl, ask=FALSE)
```



### compare deg rate between blue and red light

```{r}
# is red deg rate significantly different from blue deg rate at zt19?
hyp1 <- hypothesis(fit3ztl, hypothesis = " k_light60B  + k_light60B:zt19 >  0")

hyp1

#plot(hyp1)
```
blue is NOT different from red at ZT19

Rate at ZT 7
```{r}
# is red deg rate significantly different from blue deg rate at zt7?
hyp2 <- hypothesis(fit3ztl, hypothesis = "k_light60B < 0")

hyp2

#plot(hyp2)
```

Red faster than blue at ZT7


### calculate half-lives and compare between blue and red at ZT19

```{r}
# first get coefficients
coefs <- fixef(fit3ztl)[, "Estimate"]

#for blue
log(2) / sum(coefs[c("k_Intercept", "k_light60B", "k_zt19", "k_light60B:zt19")])

#for red
log(2) / sum(coefs[c("k_Intercept", "k_zt19")])
```

Half-lives are pretty consistent with previous calculations, nice.   

```{r}
# is red half-life significantly different from blue half-life?
hyp3 <- hypothesis(fit3ztl, hypothesis = "(log(2) / (k_Intercept + k_zt19)) > (log(2) / (k_Intercept + k_light60B + k_zt19 + k_light60B:zt19))")
hyp3

plot(hyp3)
```

Red half-life is NOT significantly lower at ZT19

### calculate half-lives and compare between blue and red at ZT7

```{r}
# first get coefficients
coefs <- fixef(fit3ztl)[, "Estimate"]

#for blue
log(2) / sum(coefs[c("k_Intercept", "k_light60B")])

#for red
log(2) / sum(coefs[c("k_Intercept")])
```

```{r}
# is red half-life significantly different from blue half-life?
hyp4 <- hypothesis(fit3ztl, hypothesis = "(log(2) / (k_Intercept)) < (log(2) / (k_Intercept + k_light60B))")
hyp4

plot(hyp4)
```

Red half-life IS different from blue at ZT7

# RVE4 

## load and check data 

```{r}
# load data
rve4 <- read_csv("CHX_RVE4_WT.csv")

# rename columns
colnames(rve4) <- c("image", "time", "zt", "light", "gt", "rep", "target", "abundance", "scaled", "zt_scaled", "norm")

# set factors, remove unnecessary columns
# also add exp_rep column to separate bio reps 
rve4 <- rve4 %>% 
  select(-c(image, norm)) %>% 
  mutate(zt = as_factor(zt),
         light = as_factor(light),
         gt = as_factor(gt),
         rep = as_factor(rep),
         target = as_factor(target))

summary(rve4)
```

```{r}
rve4 %>% 
  ggplot(aes(time, zt_scaled, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)

rve4 %>% 
  ggplot(aes(time, scaled, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)
```

```{r}
rve4 %>% 
  ggplot(aes(time, zt_scaled, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)

rve4 %>% 
  ggplot(aes(time, scaled, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)
```
make a plot with just red vs blue:

```{r}
rve4 %>% 
  filter(light != "DD", zt=="17") %>%
  ggplot(aes(time, zt_scaled, color = light, group=str_c(light,rep))) +
  scale_color_manual(values=c("60B"="blue", "60R"="red")) +
  geom_point() + 
  geom_line() +
  scale_y_log10()
```

```{r}
rve4 %>% 
  filter(light != "DD", zt=="17") %>%
  ggplot(aes(time, zt_scaled, color = light)) +
  scale_color_manual(values=c("60B"="blue", "60R"="red")) +
  geom_smooth() +
  scale_y_log10()
```


```{r}
# set up data subsets for no dark

rve4rb <- rve4 %>% filter(light != "DD") %>%
  mutate(light=relevel(light, ref="60R"))
```

with only three reps, drop the random effects...

```{r}
# no random effect for N0 or k
m3 <- bf(zt_scaled ~ N0 * exp(-k*time),
           k + N0 ~ light * zt,
           nl = TRUE)


# assign priors
priors3 <- c(prior(normal(1, 0.5), nlpar = "N0", coef = "Intercept"), #prior = 1 because that is how it was scaled
             prior(normal(0, 0.5), nlpar = "N0"), #uninformative prior for other coefficients
             prior(normal(0, 2),nlpar = "k"), #uninformative prior for rate intercept and coefficents
             prior(cauchy(0, 1), class = sigma))

# fit model 
fit3rve4 <- brm(formula = m3,
               prior = priors3,
               data = rve4rb,
               cores = 4,
               iter = 8000,
               sample_prior = "yes",
               control = list(adapt_delta = 0.95))
```

```{r}
# check model 
summary(fit3rve4)

plot(fit3rve4, ask = FALSE)
```

Rhat = 1, plots look okay.   

### compare deg rate between blue and red light


```{r}
# is red deg rate significantly different from blue deg rate at zt17?
hyp1 <- hypothesis(fit3rve4, hypothesis = "(k_light60B  + k_light60B:zt17)  = 0")

hyp1

#plot(hyp1)
```

no difference at ZT17

Rate at ZT 5
```{r}
# is red deg rate significantly different from blue deg rate at zt5?
hyp2 <- hypothesis(fit3rve4, hypothesis = "k_light60B = 0")

hyp2

#plot(hyp2)
```
no difference

### calculate half-lives and compare between blue and red at ZT17

```{r}
# first get coefficients
coefs <- fixef(fit3rve4)[, "Estimate"]

#for blue
log(2) / sum(coefs[c("k_Intercept", "k_light60B", "k_zt17", "k_light60B:zt17")])

#for red
log(2) / sum(coefs[c("k_Intercept", "k_zt17")])
```


```{r}
# is red half-life significantly different from blue half-life?
hyp3 <- hypothesis(fit3rve4, hypothesis = "(log(2) / (k_Intercept + k_zt17)) > (log(2) / (k_Intercept + k_light60B + k_zt17 + k_light60B:zt17))")
hyp3

plot(hyp3)
```

no difference

### calculate half-lives and compare between blue and red at ZT5

```{r}
# first get coefficients
coefs <- fixef(fit3rve4)[, "Estimate"]

#for blue
log(2) / sum(coefs[c("k_Intercept", "k_light60B")])

#for red
log(2) / sum(coefs[c("k_Intercept")])
```

```{r}
# is red half-life significantly different from blue half-life?
hyp4 <- hypothesis(fit3rve4, hypothesis = "(log(2) / (k_Intercept)) = (log(2) / (k_Intercept + k_light60B))")
hyp4

#plot(hyp4)
```

no difference 