---
title: "RVE Manuscript RVE4 Half-Life"
author: "Cassandra"
date: "2022-11-16"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Background

RVE4 protein degradation was assessed at ZT5 and ZT17 in monochromatic blue, monochromatic red, or constant dark light conditions.  

Julin analyzed the half-life of RVE8 from the initial four bio reps and I am trying to repeat this for RVE4 using Yuyan's experimental results.   

# Summary

In constant blue and red light, RVE4 deg rate (but not half-life) is significantly different between ZT5 and ZT17.    
In constant darkness, RVE4 half-life (and deg rate) are not significantly different between ZT5 and ZT17.     

```{r setup}
# for graphing
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())

# for Bayesian analysis 
library(posterior)
library(brms)
```

```{r}
#setwd("C:/Users/contr/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")

setwd("F:/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")
```

# RVE4

## load and check data

```{r}
# load data
rve4 <- read_csv("CHX_RVE4_WT.csv")

# rename columns
colnames(rve4) <- c("image", "time", "zt", "light", "gt", "rep", "target", "abundance", "scaled", "zt_scaled", "norm")

# set factors, remove unnecessary columns
rve4 <- rve4 %>% 
  select(-c(image, gt, norm)) %>% 
  mutate(zt = as_factor(zt),
         light = as_factor(light),
         rep = as_factor(rep),
         target = as_factor(target))

summary(rve4)
```

```{r}
rve4 %>% 
  ggplot(aes(time, zt_scaled, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)

rve4 %>% 
  ggplot(aes(time, scaled, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ light)
```

```{r}
# set up data subsets for each light quality
rve4_blue <- rve4 %>% filter(light == "60B")
rve4_red <- rve4 %>% filter(light == "60R")
rve4_dark <- rve4 %>% filter(light == "DD")
```

## Blue

### fit and compare models

Use similar analysis as for RVE8.       
**Remember that ZT5 and ZT17 were sampled here.**     

```{r cache = TRUE}
# first allow k and intercept to vary by ZT with random effect per rep 
frm1 <- bf(scaled ~ N0 * exp(-k*time),
           k + N0 ~ zt + (1|rep),
           nl = TRUE)

# determine protein abundance at time 0 
N0 <- rve4_blue %>% filter(time == 0) %>% select(scaled)
# use mean to set priors
mean(N0$scaled)
sd(N0$scaled)

# assign priors
priors1 <- c(prior(normal(0.74, 0.5), nlpar = "N0", coef = "Intercept"), #prior = mean
             prior(normal(0, 0.5), nlpar = "N0", coef = "zt17"), #uninformative prior for ZT17 being different
             prior(normal(0, 1),nlpar = "k"), #uninformative prior for rate
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "N0"),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit model 
rve4_fit1b <- brm(formula = frm1,
             prior = priors1,
             data = rve4_blue,
             cores = 4,
             iter = 24000,
             control = list(adapt_delta = 0.99))
```

```{r}
# check model 
summary(rve4_fit1b)

plot(rve4_fit1b, ask = FALSE)
```

Rhat = 1, so parameter space was sampled well. Diagnostic plots look okay.   
Not as good as RVE8 though because there are fewer reps?   

```{r cache = TRUE}
# no random effect for N0
frm2 <- bf(scaled ~ N0 * exp(-k*time),
           k ~ zt + (1|rep),
           N0 ~ zt,
           nl = TRUE)

# similar priors as before 
priors2 <- c(prior(normal(0.74, 0.5), nlpar = "N0", coef = "Intercept"), #prior = mean
             prior(normal(0, 0.5), nlpar = "N0", coef = "zt17"), #uninformative prior for ZT17 being different
             prior(normal(0, 1),nlpar = "k"), #uninformative prior for rate
             prior(cauchy(0, 1), class = sigma),
             prior(cauchy(0, 1), class = sd, nlpar = "k"))

# fit second model 
rve4_fit2b <- brm(formula = frm2,
             prior = priors2,
             data = rve4_blue,
             cores = 4,
             iter = 24000,
             control = list(adapt_delta = 0.99))
```

```{r}
# check model 
summary(rve4_fit2b)

plot(rve4_fit2b, ask = FALSE)
```

Rhat = 1, plots look okay.    

```{r cache = TRUE}
# no random effect for N0 or k
frm3 <- bf(scaled ~ N0 * exp(-k*time),
           k + N0 ~ zt,
           nl = TRUE)

# similar priors again
priors3 <- c(prior(normal(0.74, 0.5), nlpar = "N0", coef = "Intercept"), #prior = mean
             prior(normal(0, 0.5), nlpar = "N0", coef = "zt17"), #uninformative prior for ZT17 being different
             prior(normal(0, 1),nlpar = "k"), #uninformative prior for rate
             prior(cauchy(0, 1), class = sigma))

# fit third model 
rve4_fit3b <- brm(formula = frm3,
             prior = priors3,
             data = rve4_blue,
             cores = 4,
             iter = 16000,
             control = list(adapt_delta = 0.9))
```

```{r}
# check model 
summary(rve4_fit3b)

plot(rve4_fit3b, ask = FALSE)
```

Rhat = 1, plots look okay.   

```{r}
# compare the three models
LOO(rve4_fit1b, rve4_fit2b, rve4_fit3b)
```

Second model has the best fit, but the three models are comparable?    
Go with the simple third model.   

### compare deg rate between ZT5 and ZT17

```{r}
# is ZT17 deg rate significantly different from ZT5 deg rate?
hyp1 <- hypothesis(rve4_fit3b, hypothesis = "k_Intercept + k_zt17 = k_Intercept")
hyp1

plot(hyp1)

# check posterior samples of deg rate at ZT19 
k17 <- as_draws_array(rve4_fit3b, "b_k_zt17")
summarize_draws(k17)
mean(k17 <= 0) # proportion of posterior samples with k_zt19 <= 0
```

Looks like ZT17 deg rate is significantly different from ZT5 deg rate?   

### calculate half-lives and compare between ZT5 and ZT17

```{r}
# first get coefficients
coefs <- fixef(rve4_fit3b)[, "Estimate"]

#for ZT5
log(2) / coefs["k_Intercept"]

#for ZT17
log(2) / sum(coefs["k_Intercept"], coefs["k_zt17"])
```

```{r}
# is ZT17 half-life significantly different from ZT5 half-life?
hyp2 <- hypothesis(rve4_fit3b, hypothesis = "(log(2) / k_Intercept) = (log(2) / (k_Intercept + k_zt17))")
hyp2

plot(hyp2)
```

ZT17 half-life is not significantly different from ZT5 half-life with Bayesian confidence interval of 0.05.        
**Not consistent with deg rate, but error is also quite large.**        

## Red 

### fit and compare models

```{r cache = TRUE}
# update fit1b model 
rve4_fit1r <- update(rve4_fit1b, newdata = rve4_red)
```

```{r}
# check model 
summary(rve4_fit1r)

plot(rve4_fit1r, ask = FALSE)
```

Rhat = 1, so parameter space was sampled well. Diagnostic plots look okay.   

```{r cache = TRUE}
# update fit2b model
rve4_fit2r <- update(rve4_fit2b, newdata = rve4_red)
```

```{r}
# check model 
summary(rve4_fit2r)

plot(rve4_fit2r, ask = FALSE)
```

Rhat = 1, plots look okay.    

```{r cache = TRUE}
# update fit3b model
rve4_fit3r <- update(rve4_fit3b, newdata = rve4_red)
```

```{r}
# check model 
summary(rve4_fit3r)

plot(rve4_fit3r, ask = FALSE)
```

Rhat = 1, plots look okay.   

```{r}
# compare the three models
LOO(rve4_fit1r, rve4_fit2r, rve4_fit3r)
```

Third model has the best fit, so go with that.    

### compare deg rate between ZT5 and ZT17

```{r}
# is ZT17 deg rate significantly different from ZT5 deg rate?
hyp1 <- hypothesis(rve4_fit3r, hypothesis = "k_Intercept + k_zt17 = k_Intercept")
hyp1

plot(hyp1)

# check posterior samples of deg rate at ZT17 
k17 <- as_draws_array(rve4_fit3r, "b_k_zt17")
summarize_draws(k17)
mean(k17 <= 0) # proportion of posterior samples with k_zt17 <= 0
```

In constant red light, there is a significant difference between ZT5 and ZT17 deg rate?   

### calculate half-lives and compare between ZT5 and ZT17

```{r}
# first get coefficients
coefs <- fixef(rve4_fit3r)[, "Estimate"]

#for ZT5
log(2) / coefs["k_Intercept"]

#for ZT17
log(2) / sum(coefs["k_Intercept"], coefs["k_zt17"])
```

```{r}
# is ZT17 half-life significantly different from ZT5 half-life?
hyp2 <- hypothesis(rve4_fit3r, hypothesis = "(log(2) / k_Intercept) = (log(2) / (k_Intercept + k_zt17))")
hyp2

plot(hyp2)
```

ZT17 half-life is not significantly different from ZT5 half-life with Bayesian confidence interval of 0.05.        
**Also not consistent with deg rate, but error is quite large again.**    

## Dark

### fit and compare models

```{r cache = TRUE}
# update fit1b model  
rve4_fit1d <- update(rve4_fit1b, newdata = rve4_dark)
```

```{r}
# check model 
summary(rve4_fit1d)

plot(rve4_fit1d, ask = FALSE)
```

Rhat = 1, so parameter space was sampled well. Diagnostic plots look okay.   

```{r cache = TRUE}
# update fit2b model
rve4_fit2d <- update(rve4_fit2b, newdata = rve4_dark)
```

```{r}
# check model 
summary(rve4_fit2d)

plot(rve4_fit2d, ask = FALSE)
```

Rhat = 1, plots look okay.    

```{r cache = TRUE}
# update fit3b model
rve4_fit3d <- update(rve4_fit3b, newdata = rve4_dark)
```

```{r}
# check model 
summary(rve4_fit3d)

plot(rve4_fit3d, ask = FALSE)
```

Rhat = 1, plots look okay.   

```{r}
# compare the three models
LOO(rve4_fit1d, rve4_fit2d, rve4_fit3d)
```

Second model has the best fit, but all models seem comparable.     
Go with simple third model.    

### compare deg rate between ZT5 and ZT17

```{r}
# is ZT17 deg rate significantly different from ZT5 deg rate?
hyp1 <- hypothesis(rve4_fit3d, hypothesis = "k_Intercept + k_zt17 = k_Intercept")
hyp1

plot(hyp1)

# check posterior samples of deg rate at ZT17 
k17 <- as_draws_array(rve4_fit3d, "b_k_zt17")
summarize_draws(k17)
mean(k17 <= 0) # proportion of posterior samples with k_zt17 <= 0
```

In constant darkness, ZT17 deg rate is not significantly different from ZT5 deg rate.   

### calculate half-lives and compare between ZT5 and ZT17

```{r}
# first get coefficients
coefs <- fixef(rve4_fit3d)[, "Estimate"]

#for ZT5
log(2) / coefs["k_Intercept"]

#for ZT17
log(2) / sum(coefs["k_Intercept"], coefs["k_zt17"])
```

```{r}
# is ZT17 half-life significantly different from ZT5 half-life?
hyp2 <- hypothesis(rve4_fit3d, hypothesis = "(log(2) / k_Intercept) = (log(2) / (k_Intercept + k_zt17))")
hyp2

plot(hyp2)
```

ZT17 half-life is not significantly different from ZT5 half-life with Bayesian confidence interval of 0.05.        