---
title: "RVE Manuscript Protein Half-Life"
author: "Cassandra"
date: '2022-11-06'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
# for graphing
library(tidyverse)
library(RColorBrewer)
library(cowplot)
theme_set(theme_cowplot())
# for genotype manipulation
library(glue)
library(ggtext)
# for figure arrangement
library(gridExtra)

#knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)
```

```{r}
setwd("F:/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")

#setwd("C:/Users/contr/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")
```

```{r}
# set basic color scale 
simple_colors <- c("60R" = "darkred", "60B" = "darkblue", "DD" = "black")
```

# RVE8

## load and format data 

```{r}
# load data
rve8 <- read_csv("CHX_RVE8_WT.csv")

# rename columns
colnames(rve8) <- c("exp", "blot", "label", "time", "zt", "cond", "gt", "rep", "target", "abundance", "scaled", "zt_scaled")

# remove unnecessary columns and set factors
# also add column of experiment and rep 
rve8 <- rve8 %>% 
  select(-c(blot, label, gt)) %>% 
  mutate(exp = as_factor(exp),
         zt = as_factor(zt),
         cond = as_factor(cond),
         rep = as_factor(rep),
         target = as_factor(target),
         exp_rep = as_factor(paste0(exp, "_", rep)))

summary(rve8)
```

Just use the scaled ZT values for this, since this is what I had done previously.   

## rep comparison

```{r}
rve8 %>% 
  ggplot(aes(time, zt_scaled, color = exp_rep)) +
  geom_point() + 
  geom_line(aes(linetype = exp_rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ cond)
```

**Overall, reps seem pretty consistent. CHX 3 is a bit of an outlier for ZT19 60B, especially rep 1.**   

```{r}
# see what it looks like without CHX 3 for ZT19 60B
rve8 %>% filter(!(exp == "CHX 3" & zt == "19" & cond == "60B")) %>%
  ggplot(aes(time, zt_scaled, color = exp_rep)) +
  geom_point() + 
  geom_line(aes(linetype = exp_rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ cond)
```

## calculate averages

First try averages for everything. 

```{r}
rve8_all_avg <- rve8 %>% group_by(target, cond, zt, time) %>% 
  summarise(count = n(),
            avg_ab = mean(zt_scaled),
            sd = sd(zt_scaled),
            sem = sd / sqrt(count))

summary(rve8_all_avg)
```

Also calculate average excluding CHX 3 for ZT19 60B. 

```{r}
most_avg <- rve8 %>% filter(!(exp == "CHX 3" & zt == "19" & cond == "60B")) %>% 
  group_by(cond, zt, time) %>% 
  summarise(count = n(),
            avg_ab = mean(zt_scaled),
            sd = sd(zt_scaled),
            sem = sd / sqrt(count))

summary(most_avg)
```

## initial graph 

With all reps included. 

```{r}
rve8_all_avg %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10() + 
  facet_grid(. ~ zt) +
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors)
```

Excluding CHX 3 for ZT19 60B. 

```{r}
most_avg %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10() + 
  facet_grid(. ~ zt) + 
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors)
```

# RVE4

## load and format data 

```{r}
# load data
rve4 <- read_csv("CHX_RVE4_WT.csv")

# rename columns
colnames(rve4) <- c("images", "time", "zt", "cond", "gt", "rep", "target", "abundance", "scaled", "zt_scaled", "norm")

# remove unnecessary columns and set factors
rve4 <- rve4 %>% 
  select(-c(images, gt)) %>% 
  mutate(zt = as_factor(zt),
         cond = as_factor(cond),
         rep = as_factor(rep),
         target = as_factor(target))

summary(rve4)
```

## using scaled values for each ZT

### rep comparison

```{r}
rve4 %>% 
  ggplot(aes(time, zt_scaled, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ cond)
```

**Reps 1 and 2 seem pretty consistent, but rep 3 is off almost across the board. Just remove rep 3?**   

```{r}
rve4 %>% filter(rep != "3") %>% 
  ggplot(aes(time, zt_scaled, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ cond)
```

### calculate averages

First try averages for all three reps. 

```{r}
rve4_all_avg <- rve4 %>% group_by(target, cond, zt, time) %>% 
  summarise(count = n(),
            avg_ab = mean(zt_scaled),
            sd = sd(zt_scaled),
            sem = sd / sqrt(count))

summary(rve4_all_avg)
```

Also calculate average for just the first two reps. 

```{r}
two_avg <- rve4 %>% filter(rep != "3") %>% 
  group_by(cond, zt, time) %>% 
  summarise(count = n(),
            avg_ab = mean(zt_scaled),
            sd = sd(zt_scaled),
            sem = sd / sqrt(count))

summary(two_avg)
```

### initial graph 

With all three reps included. 

```{r}
rve4_all_avg %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10(breaks = c(0.1, 0.3, 1), limits = c(0.1, 1)) + 
  facet_grid(. ~ zt) +
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors)
```

With just the first two reps included. 

```{r}
two_avg %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10() + 
  facet_grid(. ~ zt) + 
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors)
```

## using just the relative quantification

### rep comparison

```{r}
rve4 %>% 
  ggplot(aes(time, abundance, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ cond)
```

Try removing rep 3 again.    

```{r}
rve4 %>% filter(rep != "3") %>% 
  ggplot(aes(time, abundance, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ cond)
```

### calculate averages

First try averages for all three reps. 

```{r}
all_avg <- rve4 %>% group_by(cond, zt, time) %>% 
  summarise(count = n(),
            avg_ab = mean(abundance),
            sd = sd(abundance),
            sem = sd / sqrt(count))

summary(all_avg)
```

Also calculate average for just the first two reps. 

```{r}
two_avg <- rve4 %>% filter(rep != "3") %>% 
  group_by(cond, zt, time) %>% 
  summarise(count = n(),
            avg_ab = mean(abundance),
            sd = sd(abundance),
            sem = sd / sqrt(count))

summary(two_avg)
```

### initial graph 

With all three reps included. 

```{r}
all_avg %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10() + 
  facet_grid(. ~ zt) + 
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors)
```

With just the first two reps included. 

```{r}
two_avg %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10() + 
  facet_grid(. ~ zt) + 
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors)
```

## using normalized abundance

I think this is what Yuyan used for her graphs. 

### rep comparison

Rep 3 wasn't normalized because the normalization sample didn't work.  

```{r}
rve4 %>% 
  ggplot(aes(time, norm, color = rep)) +
  geom_point() + 
  geom_line(aes(linetype = rep)) +
  scale_y_log10() + 
  facet_grid(zt ~ cond)
```

### calculate averages

Calculate average for just the first two reps. 

```{r}
two_avg <- rve4 %>% filter(rep != "3") %>% 
  group_by(cond, zt, time) %>% 
  summarise(count = n(),
            avg_ab = mean(norm),
            sd = sd(norm),
            sem = sd / sqrt(count))

summary(two_avg)
```

### initial graph 

With just the first two reps included. 

```{r}
two_avg %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10() + 
  facet_grid(. ~ zt) + 
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors)
```

Try to match Yuyan's graphs.  

```{r}
two_avg %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10(expand = expansion(0, 0), limits = c(0.01, 10)) + 
  facet_grid(. ~ zt) + 
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors)
```

# RVE4 and RVE8 combined

```{r}
head(rve4_all_avg)
head(rve8_all_avg)

# join RVE4 and RVE8 together
full_avg <- full_join(rve4_all_avg, rve8_all_avg)
summary(full_avg)
```

## initial graphs

```{r}
full_avg %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10() + 
  facet_wrap(target ~ zt) + 
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors)
```

Probably cleaner to just use `gridExtra` for arranging RVE4 and RVE8 together. 

### ZT arranged next to each other

```{r}
# first RVE4
rve4_rough_graph <- full_avg %>% filter(target == "RVE4") %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10(expand = expansion(0, 0),
                limits = c(0.1, 1.2)) + 
  facet_grid(target ~ zt) + 
  labs(y = "Scaled Protein Abundance",
       x = "Time After CHX Application (hours)",
       color = "Light",
       fill = "Light") + 
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors) + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        strip.text.y = element_text(face = "italic"))

# then RVE8
rve8_rough_graph <- full_avg %>% filter(target == "RVE8") %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10(expand = expansion(0, 0),
                limits = c(0.1, 1.2)) + 
  facet_grid(target ~ zt) + 
  labs(y = "Scaled Protein Abundance",
       x = "Time After CHX Application (hours)",
       color = "Light",
       fill = "Light") +
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors) +
  theme(strip.text.y = element_text(face = "italic"))

# extract the legend
light_legend <- get_legend(rve8_rough_graph)
```

```{r fig.height = 6, fig.width = 6}
layout <- rbind(c(1, 1, 1, 1),
                c(2, 2, 2, 2))

grid.arrange(rve4_rough_graph, rve8_rough_graph,
             ncol = 1, layout_matrix = layout)
```

### ZT stacked on top of each other

```{r}
# first RVE4
rve4_rough_graph <- full_avg %>% filter(target == "RVE4") %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10(expand = expansion(0, 0),
                limits = c(0.1, 1.2)) + 
  facet_grid(zt ~ target) + 
  labs(y = "Scaled Protein Abundance",
       x = "Time After CHX Application (hours)",
       color = "Light",
       fill = "Light") + 
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors) + 
  theme(strip.text.x = element_text(face = "italic"),
        legend.position = "none")

# then RVE8
rve8_rough_graph <- full_avg %>% filter(target == "RVE8") %>% 
  ggplot(aes(time, avg_ab, color = cond, fill = cond)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) +
  scale_y_log10(expand = expansion(0, 0),
                limits = c(0.1, 1.2)) + 
  facet_grid(zt ~ target) + 
  labs(y = "Scaled Protein Abundance",
       x = "Time After CHX Application (hours)",
       color = "Light",
       fill = "Light") +
  scale_color_manual(values = simple_colors) +
  scale_fill_manual(values = simple_colors) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(face = "italic"),
        legend.position = "none")
```

```{r fig.height = 4, fig.width = 8}
layout <- rbind(c(1, 1, 1, 1, 1, 2, 2, 2, 2, 3))

grid.arrange(rve4_rough_graph, rve8_rough_graph, light_legend,
             ncol = 2, layout_matrix = layout)
```

Probably just make the two graphs separately and arrange everything together in Affinity. 

# for main figure

## format data

```{r}
summary(full_avg)

# make light easier to understand 
full_avg <- full_avg %>% mutate(
  Light = as_factor(case_when(
    cond == "60B" ~ "Blue",
    cond == "60R" ~ "Red", 
    cond == "DD" ~ "Dark")),
  protein = as_factor(case_when(
    target == "RVE4" ~ "RVE4-FLAG",
    target == "RVE8" ~ "RVE8-HA"
  ))
)

# order light as a factor
full_avg$Light = as_factor(full_avg$Light) %>% 
  ordered(levels = c("Blue", "Red", "Dark"))
```

## assign graphing variables

```{r}
line_size <- 1
point_size <- 2
ribbon_alpha <- 0.25
facet_label_size = 18
axis_text_size <- 16
axis_title_size <- 18
legend_text_size <- 16
legend_title_size <- 18
legend_symbol_size <- unit(2, "line")
```

## set aesthetic scales

Set color scale for light.   

```{r}
colors <- c("Blue" = "darkblue", "Red" = "darkred", "Dark" = "black")
```

Set shapes by color. 

```{r}
shapes <- c("Red" = 1, "Blue" = 16, "Dark" = 15)
```

## graph 

```{r fig.height = 5, fig.width = 6}
# save to pdf
#pdf("RVE4_half-life.pdf", width = 6, height = 5, bg = "white")

# RVE4
full_avg %>% filter(target == "RVE4") %>% 
  ggplot(aes(time, avg_ab, color = Light, fill = Light)) + 
  geom_point(aes(shape = Light), size = point_size) + 
  geom_line(linewidth = line_size) + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = ribbon_alpha) +
  scale_y_log10(expand = expansion(0, 0),
                limits = c(0.1, 1.2)) + 
  facet_grid(zt ~ protein) + 
  labs(y = "Scaled Protein Abundance",
       x = "Time After CHX Application (hours)") + 
  background_grid(major = 'y', minor = "none") + 
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) + 
  scale_shape_manual(values = shapes) + 
  theme(axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_text_size),
        legend.text = element_text(size = legend_text_size),
        legend.title = element_text(size = legend_title_size),
        legend.key.size = legend_symbol_size,
        strip.text.x = element_text(size = facet_label_size),
        strip.text.y = element_text(size = facet_label_size))

#dev.off()

# save to pdf
#pdf("RVE8_half-life.pdf", width = 6, height = 5, bg = "white")

# RVE8
full_avg %>% filter(target == "RVE8") %>% 
  ggplot(aes(time, avg_ab, color = Light, fill = Light)) + 
  geom_point(aes(shape = Light), size = point_size) + 
  geom_line(linewidth = line_size) + 
  geom_ribbon(aes(ymin = avg_ab - sem,
              ymax = avg_ab + sem),
              linetype = 0, alpha = ribbon_alpha) +
  scale_y_log10(expand = expansion(0, 0),
                limits = c(0.1, 1.2)) + 
  facet_grid(zt ~ protein) + 
  labs(y = "Scaled Protein Abundance",
       x = "Time After CHX Application (hours)") + 
  background_grid(major = 'y', minor = "none") + 
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) + 
  scale_shape_manual(values = shapes) + 
  theme(axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_text_size),
        legend.text = element_text(size = legend_text_size),
        legend.title = element_text(size = legend_title_size),
        legend.key.size = legend_symbol_size,
        strip.text.x = element_text(size = facet_label_size),
        strip.text.y = element_text(size = facet_label_size))

#dev.off()
```

