---
title: "RVE Manuscript Protein Abundance"
author: "Cassandra"
date: "2022-12-01"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup}
# for graphing
library(tidyverse)
library(RColorBrewer)
library(cowplot)
theme_set(theme_cowplot())
# for genotype manipulation
library(glue)
library(ggtext)
# for figure arrangement
library(gridExtra)

# to check colors for accessibility 
library(colorBlindness)

#knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)
```

```{r}
setwd("F:/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")

#setwd("C:/Users/contr/OneDrive - University of California, Davis/Harmer Lab/Manuscripts/RVE/Figures/fig6_protein/")
```

```{r}
# set basic color scale 
simple_colors <- c("Red" = "darkred", "Blue" = "darkblue")
simple_light <- c("Dark" = "black", "12:12" = "grey50", "White" = "goldenrod3")
```

```{r}
# set basic shapes
simple_shapes <- c("Red" = 1, "Blue" = 16)
simple_light_shapes <- c("Dark" = 15, "12:12" = 17, "White" = 0)
```

# load and format data

```{r}
# load data
ab <- read_csv("protein_abundance.csv")

# rename columns
colnames(ab) <- c("target", "label", "zt", "cond", "rep", "abundance")

# set factors
ab <- ab %>% mutate(
  target = as_factor(target),
  label = as_factor(label),
  cond = as_factor(cond),
  rep = as_factor(rep),
  Light = as_factor(case_when(
    cond == "60B" ~ "Blue",
    cond == "60R" ~ "Red", 
    cond == "DD" ~ "Dark", 
    cond == "LL" ~ "White", 
    cond == "LD" ~ "12:12")),
  graph = as_factor(case_when(
    cond == "60B" ~ "1",
    cond == "60R" ~ "1",
    cond == "DD" ~ "2", 
    cond == "LL" ~ "2", 
    cond == "LD" ~ "2")),
  protein = as_factor(case_when(
    target == "RVE4" ~ "RVE4-FLAG",
    target == "RVE8" ~ "RVE8-HA"
  ))
)

# order RVE4 and RVE8
ab$target = as_factor(ab$target) %>% 
  ordered(levels = c("RVE4", "RVE8"))

summary(ab)
```

# rep comparison

```{r}
# RVE8
ab %>% filter(target == "RVE8") %>% 
  ggplot(aes(zt, abundance, color = rep)) + 
  geom_line() +
  geom_point() + 
  facet_wrap(~ cond)

# RVE4
ab %>% filter(target == "RVE4") %>% 
  ggplot(aes(zt, abundance, color = rep)) + 
  geom_line() +
  geom_point() + 
  facet_wrap(~ cond)
```

**RVE8 reps are pretty consistent overall. RVE4 is not quite as nice, but generally similar.**     

# calculate averages

```{r}
avg_ab <- ab %>% group_by(target, cond, Light, graph, protein, zt) %>% 
  summarise(count = n(),
            avg_ab = mean(abundance),
            sd = sd(abundance),
            sem = sd / sqrt(count))

summary(avg_ab)
```

# initial graphs

```{r}
# for blue and red
avg_ab %>% filter(cond %in% c("60B", "60R")) %>% 
  ggplot(aes(zt, avg_ab, color = Light, fill = Light)) + 
  geom_point(aes(shape = Light)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
                  ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) + 
  facet_wrap(~ target) + 
  labs(x = "ZT",
       y = "Relative Protein Abundance") + 
  scale_color_manual(values = simple_colors) + 
  scale_fill_manual(values = simple_colors) + 
  scale_shape_manual(values = simple_shapes)
```

```{r}
# for LL, LD, DD
avg_ab %>% filter(cond %in% c("LL", "LD", "DD")) %>% 
  ggplot(aes(zt, avg_ab, color = Light, fill = Light)) + 
  geom_point(aes(shape = Light)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = avg_ab - sem,
                  ymax = avg_ab + sem),
              linetype = 0, alpha = 0.2) + 
  facet_wrap(~ target) + 
  labs(x = "ZT",
       y = "Relative Protein Abundance") + 
  scale_color_manual(values = simple_light) + 
  scale_fill_manual(values = simple_light) + 
  scale_shape_manual(values = simple_light_shapes)
```

**Since the collection times were different for RVE4 and RVE8, might want to split graphs by target instead.**    

# for main figure

## assign graphing variables

```{r}
line_size <- 1
point_size <- 2
ribbon_alpha <- 0.25
facet_label_size = 18
axis_text_size <- 16
axis_title_size <- 18
legend_text_size <- 16
legend_title_size <- 18
legend_symbol_size <- unit(2, "line")
```

## set aesthetic scales

Set color scale for light.   

```{r}
colors <- c("Blue" = "darkblue", "Red" = "darkred", "Dark" = "black", 
           "12:12" = "#969696", "White" = "goldenrod3")
```

Set linetype for light.

```{r}
lines <- c("Blue" = "solid", "Red" = "solid", "Dark" = "dotted",
           "12:12" = "solid", "White" = "solid")
```

Set shapes by color. 

```{r}
shapes <- c("Red" = 1, "Blue" = 16, "Dark" = 15, "12:12" = 17, "White" = 0)
```

## graph 

```{r fig.height = 6, fig.width = 6}
# save to pdf
#pdf("RVE4_abundance.pdf", width = 6, height = 6, bg = "white")

# for RVE4
avg_ab %>% filter(target == "RVE4") %>% 
  ggplot(aes(zt, avg_ab, color = Light, fill = Light)) + 
  geom_point(aes(shape = Light), size = point_size) + 
  geom_line(aes(linetype = Light), linewidth = line_size) + 
  geom_ribbon(aes(ymin = avg_ab - sem,
                  ymax = avg_ab + sem),
              linetype = 0, alpha = ribbon_alpha) + 
  facet_grid(graph ~ protein) + 
  scale_x_continuous(breaks = seq(0, 48, by = 8)) + 
  scale_y_continuous(expand = expansion(0, 0),
                     limits = c(0, 1.05)) + 
  labs(x = "Time",
       y = "Scaled Protein Abundance") + 
  background_grid(major = 'y', minor = "none") + 
  scale_color_manual(values = colors) + 
  scale_fill_manual(values = colors) + 
  scale_shape_manual(values = shapes) + 
  scale_linetype_manual(values = lines) + 
  theme(axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_text_size),
        legend.text = element_text(size = legend_text_size),
        legend.title = element_text(size = legend_title_size),
        legend.key.size = legend_symbol_size,
        strip.text.x = element_text(size = facet_label_size),
        strip.text.y = element_blank())

#dev.off()

# save to pdf
#pdf("RVE8_abundance.pdf", width = 6, height = 6, bg = "white")

# for RVE8
avg_ab %>% filter(target == "RVE8") %>% 
  ggplot(aes(zt, avg_ab, color = Light, fill = Light)) + 
  geom_point(aes(shape = Light), size = point_size) + 
  geom_line(aes(linetype = Light), linewidth = line_size) + 
  geom_ribbon(aes(ymin = avg_ab - sem,
                  ymax = avg_ab + sem),
              linetype = 0, alpha = ribbon_alpha) + 
  facet_grid(graph ~ protein) + 
  scale_x_continuous(breaks = seq(0, 30, by = 6)) + 
  scale_y_continuous(expand = expansion(0, 0),
                     limits = c(0, 1.05)) + 
  labs(x = "Time",
       y = "Scaled Protein Abundance") + 
  background_grid(major = 'y', minor = "none") + 
  scale_color_manual(values = colors) + 
  scale_fill_manual(values = colors) + 
  scale_shape_manual(values = shapes) + 
  scale_linetype_manual(values = lines) + 
  theme(axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_text_size),
        legend.text = element_text(size = legend_text_size),
        legend.title = element_text(size = legend_title_size),
        legend.key.size = legend_symbol_size,
        strip.text.x = element_text(size = facet_label_size),
        strip.text.y = element_blank())

#dev.off()
```

**Separated RVE4 and RVE8 here because the ZT scales are different.**     

# Color Blindness Test

```{r fig.width = 9, fig.height = 9}
rve8_graph <- avg_ab %>% filter(target == "RVE8") %>% 
  ggplot(aes(zt, avg_ab, color = Light, fill = Light)) + 
  geom_point(aes(shape = Light), size = point_size) + 
  geom_line(linewidth = line_size) + 
  geom_ribbon(aes(ymin = avg_ab - sem,
                  ymax = avg_ab + sem),
              linetype = 0, alpha = ribbon_alpha) + 
  facet_grid(graph ~ protein) + 
  scale_x_continuous(breaks = seq(0, 30, by = 6)) + 
  scale_y_continuous(expand = expansion(0, 0),
                     limits = c(0, 1.05)) + 
  labs(x = "Time",
       y = "Scaled Protein Abundance") + 
  background_grid(major = 'y', minor = "none") + 
  scale_color_manual(values = colors) + 
  scale_fill_manual(values = colors) + 
  scale_shape_manual(values = shapes) + 
  theme(axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_text_size),
        legend.text = element_text(size = legend_text_size),
        legend.title = element_text(size = legend_title_size),
        legend.key.size = legend_symbol_size,
        strip.text.x = element_text(size = facet_label_size),
        strip.text.y = element_blank())

rve8_graph %>% cvdPlot()
```